---
title: "Plotly & Flexdashboard"
author: "Maliha Safdar"
date: "2025-10-21"
output: html_document
---

```{r}
library(tidyverse)
library(p8105.datasets)
library(plotly)
```

`We’re going to focus on the Airbnb data for this topic. The code below extracts what we need right now; specifically, we select only a few of the variables and filter to include a subset of the data. In part, this makes sure that the resulting dataset and plots are computationally feasible – for large datasets, you may need to downsample.`

```{r}
data(nyc_airbnb)

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(rating = review_scores_location / 2) |>
  select(
    neighbourhood_group, neighbourhood, rating, price, room_type, lat, long) |> #focus on only things that we need for the plotly
  filter(
    !is.na(rating), 
    neighbourhood_group == "Manhattan",
    room_type == "Entire home/apt",
    price %in% 100:500)
```

`We’ll use this dataset as the basis for our plots.`

### Plotly scatterplot

`There are several practical differences comparing ggplot and plot_ly, but the underlying conceptual framework is similar. We need to define a dataset, specify how variables map to plot elements, and pick a plot type.

Below we’re plotting the location (latitude and longitude) of the rentals in our dataset, and mapping price to color. We also define a new variable text_label and map that to text.

The type of plot is scatter, which has several “modes”: markers produces the same kind of plot as ggplot::geom_point, lines produces the same kind of plot as ggplot::geom_line.`


```{r}
#mode can be "markers" or "lines". Difference b/w ggplot and plotly is you can hover over and see the values and be able to zoom in. For the text_lable, when you hover over the plot, for each point it shows you price and rating. /n is used infront of Rating so that price and rating shows up on different lines.
nyc_airbnb |>
  mutate(text_label = str_c("Price: $", price, "\nRating: ", rating)) |> 
  plot_ly(
    x = ~lat, y = ~long, type = "scatter", mode = "markers",
    color = ~price, text = ~text_label, alpha = 0.5)
```

`This can be a useful way to show the data – it gives additional information on hovering and allows you to zoom in or out, for example.`

### Plotly boxplot

`Next up is the boxplot. The process for creating the boxplot is similar to above: define the dataset, specify the mappings, pick a plot type. Here the type is box, and there aren’t modes to choose from.`

```{r}
nyc_airbnb |> 
  mutate(neighbourhood = fct_reorder(neighbourhood, price)) |> 
  plot_ly(y = ~price, color = ~neighbourhood, type = "box", colors = "viridis")
```

`Again, this can be helpful – we have a five-number summary when we hover, and by clicking we can select groups we want to include or exclude.`

### Plotly barchart

`Lastly, we’ll make a bar chart. Plotly expects data in a specific format for bar charts, so we use count to get the number of rentals in each neighborhood (i.e. to get the bar height). Otherwise, the process should seem pretty familiar …`


```{r}
nyc_airbnb |> 
  count(neighbourhood) |> 
  mutate(neighbourhood = fct_reorder(neighbourhood, n)) |> 
  plot_ly(x = ~neighbourhood, y = ~n, color = ~neighbourhood, type = "bar", colors = "viridis")
```

`Interactivity in bar charts is kinda neat, but needs a bit more justification – you can zoom, which helps in some cases, or you could build in some addition information in hover text.`

### ggplotly

`You can convert a ggplot object straight to an interactive graphic using ggplotly.

For example, the code below recreates our scatterplot using ggplot followed by ggplotly`

```{r}
scatter_ggplot = 
  nyc_airbnb |>
  ggplot(aes(x = lat, y = long, color = price)) +
  geom_point(alpha = 0.25) +
  coord_cartesian()

ggplotly(scatter_ggplot)
```

`We can recreate our boxplot in a similar way.`


```{r}
box_ggplot = 
  nyc_airbnb |> 
  mutate(neighbourhood = fct_reorder(neighbourhood, price)) |> 
  ggplot(aes(x = neighbourhood, y = price, fill = neighbourhood)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplotly(box_ggplot)
```

`If I really want an interactive plot to look good, I’ll use plot_ly to build it – ggplot was designed with static plots in mind, and the formatting and behavior of ggplotly is less visually appealing (to me) than plot_ly.

I use ggplot for static plots, and I make static plots way, way more frequently than interactive plots. Sometimes I’ll use ggplotly on top of that for some quick interactivity; this can be handy to do some zooming or inspect outlying features.`


### flexdashboard


